---
title: "Code Book. Getting and cleaning data. Course Project."
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This codebook describes the steps taken in the file "run_analysis". 

### Download data from zip link.

* We create temp file that we use to download our zip link.  
```{r temp}
temp <- tempfile()
```
* Download archive into temp file.   
```{r download, echo=TRUE,cache=FALSE}
download.file("https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip",temp,method = "curl")

```
* Create temporary directory temp1, in which we will unzip our archive.  
```{r temp1}
temp1<-tempdir()

```
* Unzip our archive (temp) into the temporary directory (temp1)  
```{r unzip}
unzip(zipfile = temp,exdir = temp1)
```
* Show all files and directories in the temp1  
```{r list_all}
list.files(temp1)
list.dirs(temp1)
```

### Extracting test data 

* Download packages dplyr and tidyr

```{r install packages}
##install.packages("dplyr")
##install.packages("tidyr")
require(dplyr)
require(tidyr)
```

* Using readtable function to download test data, features, labels and subject data

```{r test}
features<-read.table(paste(temp1,"/UCI HAR Dataset/features.txt",sep=""))
alabels<-read.table(paste(temp1,"/UCI HAR Dataset/activity_labels.txt", sep = ""))
x_test<-read.table(paste(temp1,"/UCI HAR Dataset/test/X_test.txt",sep=""))
y_test<-read.table(paste(temp1, "/UCI HAR Dataset/test/y_test.txt",sep = ""))
subject_test<-read.table(paste(temp1, "/UCI HAR Dataset/test/subject_test.txt",sep = ""))

```

* Cleaning test data. The final data set (test) has all information about the subject and object of test observations.

```{r clean1}
x_test<-tbl_df(x_test)
y_test<-tbl_df(y_test) %>%
  full_join(alabels,by="V1") %>%
  rename(id=V1,label=V2) %>%
  mutate(sample=rep("test",2947))
test<-tbl_df(bind_cols(x_test,y_test))
test$subject<-subject_test$V1
str(test)
```

### Extracting test data

* Using readtable function to download train data and subject.
```{r train}
x_train<-read.table(paste(temp1,"/UCI HAR Dataset/train/X_train.txt",sep=""))
y_train<-read.table(paste(temp1,"/UCI HAR Dataset/train/y_train.txt",sep=""))
subject_train<-read.table(paste(temp1,"/UCI HAR Dataset/train/subject_train.txt",sep=""))

```

### Cleaning data

* Cleaning train data. The final data set (train) has all information about the subject and object of test observations.
```{r clean2}
x_train<-tbl_df(x_train)
y_train<-tbl_df(y_train) %>%
  full_join(alabels,by="V1") %>%
  rename(id=V1,label=V2) %>%
  mutate(sample=rep("train",7352))
train<-tbl_df(bind_cols(x_train,y_train))
train$subject<-subject_train$V1
str(train)
```
* Check the number of rows in two data sets.
```{r number}
b<-nrow(test)+nrow(train)
b
```

* Merge train and test data sets. The "df" dataframe are clean and tidy.

```{r merge}
df<-bind_rows(train,test)
names<-make.unique(as.character(features$V2), sep = "_")
colnames(df)<-c(names,"label","activity","sample","subject")
dim(df)
glimpse(df[,560:565])
head(df[,560:565])
```

### Extracting mean and stdev 

* Extracting only the measurements on the mean and standard deviation for each measurement.

```{r}
df1<-df %>%
  select(matches("mean()|std()"),label, activity, sample, subject)
head(df1[,85:90])
```

### Summarize average values

* This script creates another independent tidy data set with the average of each variable for each activity and each subject.
```{r}
df1<-tbl_df(df1)
df2<-df1 %>%
  gather(name,value,-activity,-sample,-subject,-label) %>%
  group_by(activity,subject) %>%
  summarise(average=mean(value))

print(head(df2,10))

```

